Current Sheet Theory

### 2024-06-03 index.qmd
fix typo
chores
update motivations

---
title: "Current Sheet Theory"
code-links:
  - text: Tana
    href: https://app.tana.inc?nodeid=OzzX0sVA4_6X
  - text: Zotero
    href: zotero://select/library/collections/DZHNUFNV
---

@@

@@

$$
u_{2x} := λ_x u_{1x}, \quad u_{2y} := λ_y u_{1y}, \quad u_{2z} := λ_z u_{1z}, \quad n_2 := λ_n n_1
$$

$$
Λ_x := 1 + λ_x λ_n, \quad Λ_y := 1 + λ_y λ_n, \quad Λ_z := 1 + λ_z λ_n
$$

@@

The Ampere's law (@eq-Jx) become:

$$
B_0 \sin θ θ' = Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
$$$$
\Rightarrow 
θ' = \mp \frac{Λ_y n_1 u_1 }{B_0} + \frac{Λ_z Γ_1}{B_z}
$${#eq-theta}

@@

The system $S$ could be fully determined by $λ_n, λ_z, c, B_0$ parameters, provided $n(z)$ profiles.

$$S=S(z; n(z); λ_n, λ_z, c, B_0)$$

>>>>

---
title: "Current Sheet Theory"
format:
  html: default
  pptx:
    shift-heading-level-by: -1
code-links:
  - text: Tana
    href: https://app.tana.inc?nodeid=OzzX0sVA4_6X
  - text: Zotero
    href: zotero://select/library/collections/DZHNUFNV
---

@@

## Motivation

In MHD theory, for rotational rotational discontinuity, we have this following relationship

$$
[[\vec{U}]]=\left(\frac{\xi \rho}{\mu_0}\right)^{1 / 2}\left[\left[\frac{\vec{B}}{\rho}\right]\right]
$$

where $\xi \equiv 1-\frac{p_{\parallel}-p_{\perp}}{B^2 / \mu_0}$.

However, observations of solar wind discontinuities reveals discrepancies between

- Alfven velocity and plasma velocity change across discontinuities
- Anisotropic MHD theory-predicted and directly measured ion anisotropies

Multifluid theory is proposed to address these discrepancies. The basic idea is that we could have zero bulk velocity with non zero pressure.

### Previous Work

On the extreme of multiple fluid theory, we have the kinetic Harris model for the current sheet.

> Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005].

@steinhauerMultifluidModelOnedimensional2008 built a multifluid model to study the steady state magnetotail current sheet.

![Ion trajectories in model magnetic field.](figures/jgra18999-fig-0001.png)

For the symmetric three-fluid system with $B_y=0$,

![Magnetic field, current density, ion density and out-of-plane flow for Bz = 0.1, βe = 0.5, βi = 1. Solid lines are fluid model results and dot symbols are from the Harris sheet solution.](figures/jgra18999-fig-0003.png)

Here, we are interested in the case of a force-free current sheet and start from the simplest case of two ion species.

@@

$$
u_{αx} := λ_{αx} u_{1x}, \quad u_{αy} := λ_{αy} u_{1y}, \quad u_{αz} := λ_{αz} u_{1z}, \quad n_α := λ_{αn} n_1
$$

$$
Λ_x := 1 + \sum λ_{αx} λ_{αn}, \quad Λ_y := 1 + \sum λ_{αy} λ_{αn}, \quad Λ_z := \sum 1 + λ_{αz} λ_{αn}
$$

@@

The Ampere's law (@eq-Jy) become:

$$
- B_0 \sin θ θ' = Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
$$

$$
\Rightarrow 
θ' = \mp \frac{Λ_y n_1 u_1 }{B_0} + \frac{Λ_z Γ_1}{B_z}
$${#eq-theta}

@@

The system $S$ could be fully determined by $λ_n, λ_z, B_0$ parameters, provided $n(z)$ profiles.

$$
S=S(z; n(z); λ_n, λ_z, B_0)
$$
